---
title: "Bibliography of the INLPO"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    css: extra.css
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Bibliography of the INLPO}
  %\VignetteEncoding{UTF-8}
---

```{r "setup", include=FALSE}
library(inlpubs)
cit <- inlpubs::pubs$citation
```

```{r "bib_dat", include=FALSE}
key <- rownames(pubs)

cite <- apply(pubs, 1, function(x) {
  cit <- format(x$citation$author, "family")
  if (length(cit) > 2) {
    cit <- sprintf("%s and others", cit[1])
  } else if (length(cit) == 2) {
    cit <- paste(cit, collapse = " and ")
  }
  cit
})

badge <- unlist(apply(pubs, 1, function(x) {
  doi <- x$citation$doi
  fmt <- "<a class='pubs-tag %s' href='%s' target='_blank'>DOI&nbsp;%s</a>"
  if (is.null(doi)) {
    sprintf(fmt, "other", "https://www.doi.org/", "not available")
  } else {
    sprintf(fmt, "doi", sprintf("https://doi.org/%s", doi), doi)
  }
}))

title <- apply(pubs, 1, function(x) x$citation$title)

author <- apply(pubs, 1, function(x) {
  aut <- format(x$citation$author, c("given", "family"))
  for (i in seq_along(aut)) {
    orcid <- x$citation$author[i]$comment["ORCID"]
    if (!is.null(orcid)) {
      href <- sprintf("https://orcid.org/%s", orcid)
      icon <- sprintf("<a href='%s' target='_blank'>%s</a>", href, inlpubs:::orcid_icon)
      aut[i] <- paste0(aut[i], icon)
    }
    email <- x$citation$author[i]$email
    if (!is.null(email)) {
      href <- sprintf("mailto: %s", email)
      icon <- sprintf("<a href='%s'>%s</a>", href, inlpubs:::email_icon)
      aut[i] <- paste0(aut[i], icon)
    }
  }
  n <- length(aut)
  if (n > 2) aut[-n] <- paste0(aut[-n], ",")
  if (n > 1) aut[n] <- paste("and", aut[n])
  paste(aut, collapse = " ")
})

reference <- vapply(cit, function(x) attr(unclass(x)[[1]], "textVersion"), character(1))

bib <- lapply(cit, function(x) {
  b <- utils::toBibtex(x)
  idx <- c(1, length(b))
  out <- stringi::stri_wrap(b[-idx], width = 60, simplify = FALSE, indent = 2, exdent = 4)
  b[-idx] <- vapply(out, function(y) {
    stringi::stri_c(y, collapse = "\n", ignore_null = TRUE)
  }, character(1))
  b
})
```

This document contains the bibliography for published reports, journal articles, maps, and theses related to scientific monitoring and research conducted by the U.S. Geological Survey (USGS),Idaho Water Science Center, Idaho National Laboratory Project Office ([INLPO]). The bibliography includes entries for `r nrow(inlpubs::pubs)` publications published during `r min(pubs$year)` through `r max(pubs$year)`. Each entry contains the digital object identifier ([DOI]), title, name(s) of individual [authors], a text representation of the reference to be employed for printing, a BibTeX entry for [LaTeX] users, and abstract. The entry may also include an annotation if no abstract is available. Hyperlinks to the [ORCiD] identifier and (or) email address may be located to the right of the authors name. The arrangement of the entries is by year of publication, in descending order, and subordinately by authors, in the alphabetical order of their names.

---

```{r "bib_lst", echo=FALSE, results="asis"}
for (yr in seq(max(pubs$year), min(pubs$year))) {
  cat(sprintf("## %s", yr), "\n\n")
  idx <- which(pubs$year %in% yr)
  if (length(idx) == 0) cat("<p>None</p>", "\n\n")
  for (i in idx) {
    cat(sprintf("### %s {#%s}", cite[i], key[i]), "\n\n")
    if (!is.null(badge[i])) {
      cat("<div style='paddding:10px 0'>", "\n")
      cat(badge[i], "\n")
      cat("</div>", "\n\n")
    }
    cat("#### Title", "\n\n")
    cat(title[i], "\n\n")
    cat("#### Authors", "\n\n")
    cat(author[i], "\n\n")
    cat("#### Suggested citation", "\n\n")
    cat(reference[i], "\n\n")

    cat("<details>", "\n")
    cat("<summary>BibTeX citation</summary>", "\n")
    cat("```", "\n")
    cat(bib[[i]], sep = "\n")
    cat("```", "\n")
    cat("</details>", "\n")
    x <- pubs[i, "abstract"]
    if (!is.na(x) && nzchar(x)) {
      cat("<details>", "\n\n")
      cat("<summary>Abstract</summary>", "\n")
      cat("\n")
      cat(stringi::stri_unescape_unicode(x), sep = "\n")
      cat("\n")
      cat("</details>", "\n\n")
    }
    x <- pubs[i, "annotation"]
    if (!is.na(x) && nzchar(x)) {
      cat("<details>", "\n")
      cat("<summary>Annotation</summary>", "\n")
      cat("\n")
      cat(stringi::stri_unescape_unicode(x), sep = "\n")
      cat("\n")
      cat("</details>", "\n\n")
    }
  }
}
```

<!-- Embedded References -->

[INLPO]: https://www.usgs.gov/centers/idaho-water-science-center/science/idaho-national-laboratory-project-office
[DOI]: https://www.doi.org/
[authors]: inl-authors.html
[LaTeX]: https://www.latex-project.org/
[ORCiD]: https://orcid.org/
