---
title: "Analysis of Publications in the INLPO"
output:
  rmarkdown::html_document:
    toc: false
    css: extra.css
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Analysis of Publications in the INLPO}
  %\VignetteEncoding{UTF-8}
---

```{r "setup", include=FALSE}
library(inlpubs)
cit <- inlpubs::pubs$citation
dt_opt <- list(
  "dom" = "Blfrtip",
  "pagingType" = "full",
  "lengthMenu" = list(
    c(10L, 25L, 50L, -1L),
    c("10", "25", "50", "all")
  ),
  "buttons" = list(
    list(
      "extend" = "copyHtml5",
      "titleAttr" = "Copy visible data (tab deliminated) to clipboard",
      "title" = NULL,
      "exportOptions" = list(
        "columns" =  ":visible",
        "rows" = ":visible"
      )
    ),
    list(
      "extend" = "csvHtml5",
      "titleAttr" = "Save all rows in a CSV (comma deliminated) file format",
      "filename" = "inlpubs-text-mining"
    )
  )
)
```

```{r "bib_stats", echo=FALSE, results="asis"}
institutions <- vapply(cit, function(x) {
  desc <- attr(unclass(x)[[1]], "bibtype")
  if (is.null(x$institution)) character(1) else x$institution
}, character(1))
is <- institutions %in% "U.S. Geological Survey"
percent_usgs <- (sum(is) / length(is) * 100) |> round(digits = 1) |> format(nsmall = 1)
years <- table(pubs$year)
mean_years <- mean(years) |> round(digits = 1) |> format(nsmall = 1)
sd_years <- sd(years) |> round(digits = 1) |> format(nsmall = 1)
```

Idaho National Laboratory Project Office ([INLPO]) publication types include journal articles, conference proceedings, university theses, U.S. Geological Survey (USGS) Series, U.S. Atomic Energy Commission reports, and Idaho Department of Water Administration reports. The percentage frequency distribution of INLPO publication types is provided in table 1. And the number of citations per year is shown in figure 1. USGS-Series publications compose the largest percentage of the INLPO bibliography at `r percent_usgs` percent. The mean number of publications per year is `r mean_years`, with a standard deviation of `r sd_years`.

<b>Table 1.</b> &nbsp; Type and number of INL Project Office publications (`r  min(pubs$year)`--`r  max(pubs$year)`).
<div class='subtitle'>
[<b>Type of publication</b>: the period of time over which they were published are enclosed within parentheses.
 <b>Count</b>: number of publications.
 <b>Percent</b>: is the percentage of the total number of publications.
   Note that percentages may not sum to 100 percent due to rounding.]
</div>

```{r "ref_typ_tbl", echo=FALSE, results="asis"}
m <- do.call("rbind", lapply(cit, function(x) {
  desc <- attr(unclass(x)[[1]], "bibtype")
  type <- if (is.null(x$type)) character(1) else x$type
  institution <- if (is.null(x$institution)) "" else x$institution
  if (desc == "TechReport") desc <- institution
  c(desc, type)
}))
d <- data.frame(m[!duplicated(m), , drop = FALSE], stringsAsFactors = FALSE)
d <- d[order(d[, 1], d[, 2]), ]
is <- d[, 2] == ""
d[is, 2] <- d[is, 1]
d[is, 1] <- NA
d[, 3] <- as.integer(table(m)[d[, 2]])
d <- d[order(d[, 1], d[, 3], decreasing = TRUE, na.last = FALSE), ]
ran <- vapply(seq_len(nrow(d)), function(i) {
  d[i, is.na(d[i, 1]) + 1L]
  is <- m[, 1] == d[i, 2] | m[, 2] == d[i, 2]
  ran <- range(pubs$year[is])
  if (identical(ran[1], ran[2])) format(ran[1]) else paste(ran, collapse = "--")
}, character(1))
d[d[, 2] == "Article", 2] <- "Article from a journal"
d[d[, 2] == "InProceedings", 2] <- "Article in a conference proceedings"
d[d[, 2] == "Proceedings", 2] <- "Proceedings of a conference"
d[d[, 2] == "MastersThesis", 2] <- "Master's thesis"
d[, 2] <- sprintf(
  "%s <span style='color:#808080;font-size:small;'>(%s)</span>",
  d[, 2], ran
)
d[, 4] <- d[, 3] / sum(d[, 3]) * 100
d[, 4] <- formatC(round(d[, 4], digits = 1), format = "f", digits = 1)
nm <- as.character(na.omit(unique(d[, 1])))
grp <- lapply(nm, function(x) range(which(d[, 1] %in% x)))
names(grp) <- nm
d <- rbind(d, c("", "Total", sum(d[, 3]), ""))
rownames(d) <- NULL
d[, 1] <- NULL
x <- knitr::kable(d,
  format = "html",
  col.names = c("Type of publication", "Count", "Percent"),
  align = c("l", "r", "r"), escape = FALSE
)
x <- kableExtra::kable_styling(x, c("row-border", "condensed"),
  full_width = FALSE, position = "left"
)
for (i in seq_along(grp)) {
  x <- kableExtra::pack_rows(x, names(grp)[i], grp[[i]][1], grp[[i]][2])
}
kableExtra::row_spec(x, row = nrow(d), italic = TRUE)
```

```{r "graph_prod", echo=FALSE, results="asis", fig.height=4, fig.cap=sprintf("<b>Figure 1.</b> &nbsp; Distribution of INL Project Office publications (%s--%s). USGS and non-USGS publication types are distinguished by color.", min(pubs$year), max(pubs$year))}

tbl <- table(pubs$year)
d1 <- data.frame(
  x = as.Date(sprintf("%s-01-01", names(tbl))), y1 = as.integer(tbl),
  y2 = NA_integer_, stringsAsFactors = FALSE
)
is <- vapply(pubs$citation, function(x) {
  !is.null(x$institution) && x$institution %in% "U.S. Geological Survey"
}, logical(1))
tbl <- table(pubs$year[is])
d2 <- data.frame(
  x = as.Date(sprintf("%s-01-01", names(tbl))), y2 = as.integer(tbl),
  stringsAsFactors = FALSE
)
d1$y2[match(d2$x, d1$x)] <- d2$y2
ylab <- "Total number of publications per year"
cols <- c("Non-USGS" = "#B47846", "USGS" = "#4682B4")

npubs <- apply(d1[, -1], 1, sum, na.rm = TRUE)
ylim <- range(pretty(0:max(npubs)))

dates <- d1[, 1]
pretty_dates <- pretty(dates)
is <- dates %in% pretty_dates
dates[!is] <- NA
pretty_dates <- dates[is]
at <- which(is)

graphics::par(mar = c(2.1, 3.1, 1.1, 0.1))

graphics::barplot(
  height = t(as.matrix(d1[, -1])),
  names.arg = dates,
  ylab = ylab,
  col = cols,
  space = 0,
  border = NA,
  ylim = ylim,
  yaxt = "n",
  xaxt = "n",
  cex.axis = 0.7,
  cex.lab = 0.7,
  mgp = c(2, 0.5, 0)
)

graphics::axis(
  side = 1,
  at = at,
  labels = format(pretty_dates),
  tck = 0.02,
  lwd = 0,
  lwd.ticks = 0.5,
  cex.axis = 0.7,
  mgp = c(2, 0.5, 0)
)

graphics::axis(
  side = 2,
  ylab = ylab,
  las = 2,
  tck = 0.02,
  lwd = 0,
  lwd.ticks = 0.5,
  cex.axis = 0.7,
  mgp = c(2, 0.5, 0)
)

graphics::axis(
  side = 3,
  at = at,
  labels = NA,
  tck = 0.02,
  lwd = 0,
  lwd.ticks = 0.5,
  mgp = c(2, 0.5, 0)
)

graphics::axis(
  side = 4,
  labels = NA,
  las = 2,
  tck = 0.02,
  lwd = 0,
  lwd.ticks = 0.5,
  mgp = c(2, 0.5, 0)
)

graphics::legend(
  x = "topright",
  legend = rev(names(cols)),
  fill = rev(cols),
  border = NA,
  inset = c(0.02, 0.05),
  cex = 0.7,
  box.lty = 1,
  box.lwd = 0.5,
  xpd = NA,
  bg = "#FFFFFFE7"
)

graphics:: box(lwd = 0.5)
```

```{r "tm_data", echo=FALSE, results="asis"}
m <- mine_text(pubs, lowfreq = 3L)
tm_dat <- data.frame("term" = rownames(m), "count" = rowSums(m), "support" = rowSums(m > 0))
top_word <- stringi::stri_trans_totitle(tm_dat[1, "term"])
top_count <- format(tm_dat[1, "count"], big.mark = ",")
```

A total of `r format(nrow(tm_dat), big.mark = ",")` unique words were identified in the titles and abstracts of INLPO publications. A frequency count of these words is provided in table 2 and sorted in order of decreasing frequency. `r top_word` is the most used word with a count of `r top_count`.

<b>Table 2.</b> &nbsp; The frequency of words used in titles and abstracts
(`r  min(pubs$year)`--`r  max(pubs$year)`).
Words with a frequency count of less than three were excluded from the table.
<div class='subtitle'>
[<b>Word</b>: is a word taken from the unstructured text.
 <b>Frequency count</b>: is the number of appearances in the documents.
 <b>No. of publications</b>: is the number of documents containing the word.]
</div>

```{r "tm_tbl", echo=FALSE, results="asis"}
tbl <- DT::datatable(tm_dat,
  options = dt_opt,
  extensions = "Buttons",
  class = "hover row-border compact",
  colnames = c("Word", "Frequency count", "No. of publications"),
  rownames = FALSE,
  escape = FALSE,
  width = 600,
  elementId = "htmlwidget-tm-tbl"
)
tbl <- DT::formatCurrency(tbl, "count", currency = "", digits = 0)
tbl <- DT::formatCurrency(tbl, "support", currency = "", digits = 0)
tbl
```

<!-- Embedded References -->

[INLPO]: https://www.usgs.gov/centers/idaho-water-science-center/science/idaho-national-laboratory-project-office
